Dirac neutrino mass generation from Majorana messenger

### 2019-08-12 draft.tex
Updates from Overleaf

% \subsection{Scalar mass spectrum}
%\label{sec:ScaMassSpect}
To establish the scalar spectrum we expand the scalar fields as
%
\begin{align*}
  H = \begin{pmatrix}G^+ \\ \frac{1}{\sqrt{2}} (\varphi+v_H+iG) \end{pmatrix} \,,&\hspace{1cm}
  \eta = \begin{pmatrix}\eta^{+} \\ \frac{1}{\sqrt{2}}(\eta_R+i\eta_I) \end{pmatrix} \,,\\
  S = \frac{1}{\sqrt{2}} (S_R+v_{S}+iS_I)\,,&\hspace{1cm}\sigma = \frac{1}{\sqrt{2}} (\sigma_R+i\sigma_I),
\end{align*}
%
with $v= 246.22$ GeV.  
By minimizing the scalar potential we find that the mass-squared parameters $\mu_H^2$ and $\mu_S^2$
%
\begin{align*}
    \mu^{2}_{H} =& -\frac{1}{2}\left( \lambda_{H}  v^2 + \lambda_{1}  v_S\right)\,,  &
    \mu^{2}_{S} =& -\frac{1}{2}\left( \lambda_{S}  v^{2}_{S} + \lambda_{1}  v^2\right)\,.
\end{align*}
%
Of the original twelve scalar degrees of freedom in the model, $W^{\pm}$, $Z^{0}$ and $Z^{\prime}$ absorve four of them (the Goldstone bosons $G^{\pm}$,$G$ and $S_{I}$). There is a physical neutral CP-even Higgs state ($\varphi$), one scalar state ($S_{R}$), two CP-even scalar states ($\sigma_{R}$ and $\eta_{R}$), two CP-odd scalar states ($\sigma_{I}$ and $\eta_{I}$) and one charged scalar ($\eta^{\pm}$). The mass spectrum for the scalar sector is given by
\begin{align*}
    m_{\eta^{\pm}}^{2} &= \mu_{\eta}^{2} + \frac{1}{2} (\lambda_{3} \upsilon^{2}_{H} + \lambda_{5} \upsilon^{2}_{s} )\,, \\
    m_{H}^{2} &= \lambda_{H} \upsilon_{H}^{2} + \mathcal{O}()\,, \\
    m_{S}^{2} &= \lambda_{S} \upsilon_{S}^{2} + \mathcal{O}()\,.
\end{align*}
%
The neutral mass eigenstate are defined as
%
\begin{align*}
    \begin{pmatrix}\chi_{(R,I)_{1}} \\ \chi_{(R,I)_{2}} \end{pmatrix} =& \begin{pmatrix} c_{\theta_{(R,I)}} & -s_{\theta_{(R,I)}} \\ s_{\theta_{(R,I)}} & c_{\theta_{(R,I)}} \end{pmatrix} \begin{pmatrix}\sigma_{(R,I)} \\ \eta_{(R,I)} \end{pmatrix} \,,
\end{align*}
%
where $\tan2\theta_{(R,I)} = \pm2c/(b-a)$ and we defined $a = m^{2}_{\eta^{\pm}} + \frac{1}{2}\lambda_{7} \upsilon^{2}_{H}$, $b = \mu^{2}_{\sigma}+\frac{1}{2}(\lambda_{2} \upsilon^{2}_{H} + \lambda_{4} \upsilon^{2}_{S})$ and $c = \frac{1}{2}\mu \upsilon_{H} \upsilon_{S}$\,. The corresponding mass eigenvalues are
\begin{align*}
    m_{\chi_{R(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} + 4c^{2}} \right)\,, \\
     m_{\chi_{I(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} - 4c^{2}} \right)\,,
\end{align*}

On the other hand, we assume that the heavy Majorana fermions are already in the diagonal basis defined as $
\begin{pmatrix}
  \psi_1 &
  \psi_2 &
  \psi_3
\end{pmatrix}^{\operatorname{T}}
$, with masses $M_{\psi_a}$ ($a=1,2,3$).
\section{Neutrinos}
\label{sec:Neutrinos}
%
\begin{figure}
\centering
\includegraphics[scale=0.75]{Neutrino_Loop.pdf}
\caption{One-loop generation of Dirac neutrino mass}
\label{fig:zee}
\end{figure}
%
 The Dirac neutrino mass matrix in this model is generated at one loop-level according to the diagram in  fig.~\ref{fig:zee}. The expression for $\mathcal{M}_{\nu}$ is given by~\cite{Reig:2018mdk}
%
\begin{align}
(\mathcal{M}_{\nu})_{ij} = \frac{\lambda_8 v_S v}{32 \pi^{2}} \sum_{a=1}^{3} \frac{h_{i a} M_{\psi_{a}}y^{\dagger}_{a j}} {m_{\chi_{R_2}}^{2}-m_{\chi_{R_1}}^{2}} \left[ F\left( \frac{m_{\chi_{R_2}}^{2}}{M_{\psi_{a}}} \right) - F\left( \frac{m_{\chi_{R_1}}^{2}}{M_{\psi_{a}}} \right) \right] + (R \to I)\,,
\end{align}
%
%check the following formula:
where $F(x) =x \log x/(x-1)$ is the loop function, $M_{\psi_{a}}$ is
the fermion mass and $\chi_{R_{(1,2)}}$ ( $\chi_{I_{(1,2)}}$ ) are the
two CP-even (CP-odd) mass eigenstates.
To obtain neutrino masses it is necessary that $\mu \neq 0$. In this
way, in addition to the suppression from the seesaw mechanism, we can
have further suppression in the neutrino mass matrix for small values
of $\mu$.
The structure of the effective neutrino mass matrix is given by the
product $(M_{\nu})_{ij} \propto h_{i a} y_{a j}$, which is similar to
the structure of the neutrino mass matrix for the tree-level seesaw
mechanism for Dirac neutrinos~\cite{Chulia:2016ngi}. If only one
fermion $\psi_R$ is added, we will have two massless neutrinos, which
does not agree with the current data obtained from the neutrino
experiments~\cite{deSalas:2017kay}. In our case, we will assume the
existence of three fermions, generating Dirac scotogenic masses for
the two neutrinos ($\nu_{k}$ is massless due to the assignment of
charges).

The mass matrix for Dirac neutrinos is diagonalized by
\begin{align}
(U^{\nu}_R)^{\dagger} \mathcal{M}_{\nu} U^{\nu}_L = \mathcal{M}_{\nu}^{\text{diag}} = \text{diag}(m_1, m_2, m_3)\,,
\end{align}
where $U^{(\nu)}_R$ and $U^{(\nu)}_L$ are unitary $3 \times 3$
matrices. The leptonic charged currents with non-massless neutrinos
are parametrized in terms of the unitary mass
\begin{align}
U_{\text{PMNS}} = \left( U^{\ell}_L \right)^{\dagger} U^{\nu}_L\,,
\end{align}
where is the left diagonalization matrix of the charged leptons.
As we already are in the diagonal basis after the calculation of the
one-loop neutrino masses, we can set
$U_{L}^{\ell} = \mathds{1}$. Moreover, we assume that $U_{R}^{\nu} = \mathds{1}$.
Therefore, we can express the Dirac neutrino mass matrix in the
following way
\begin{align}
(\mathcal{M}_{\nu})_{ij} = (U_{\text{PMNS}})_{ij} (\mathcal{M}_{\nu}^{\text{diag}})_{j}\,,
\end{align}
where the matrix $U_{\text{PMNS}}$ is expressed in the standard parameterization as
\begin{align}
U_{\text{PMNS}} = \begin{pmatrix}
    1 & 0		& 0 \\
    0 & c_{23}  & s_{23} \\
    0 & -s_{23} & c_{23}
\end{pmatrix}
\begin{pmatrix}
    c_{13}  &  0 & s_{13}e^{-i\delta} \\
    0 		&  1 & 0 \\
    -s_{13}e^{i\delta} &  0 & c_{13}
\end{pmatrix}
\begin{pmatrix}
    c_{12}  & s_{12} & 0 \\
    -s_{12} & c_{12} & 0 \\
    0		& 0		 & 1
\end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin \theta_{ij}$ and $c_{ij} = \cos \theta_{ij}$ and $\theta_{ij}$ are neutrino mixing angles and $\delta$ is CP-violating phase.

@@

For $i \neq j$ we have CLFV processes, these processes are severely restricted by the current bounds , which set limits to the Yukawas. The most restrictive experimental bounds is decay $\mu \to e \gamma$, which is several orders of magnitude greater than radiative decays $\tau \to~l\gamma$, with $l = e, \mu$. These restrictions are given by Br$(\mu \to e\gamma)$ \textless $5.3 \times 10^{-13} $~\cite{Adam:2013mnn}, Br$(\tau \to e\gamma) $\textless$ 3.3 \times 10^{-8}$ and Br$(\tau \to~\mu\gamma) $\textless$ 4.4 \times 10^{-8}$~\cite{Aubert:2009ag, Bona:2007qt, Miyazaki:2012mx}.

For $i = j$ we have processes that contribute to the leptonic magnetic dipole moment. In the case of the anomalous muon magnetic dipole moment, the experimental measurements does not agree with the values predicted by the SM~\cite{Lindner:2016bgg}. This can be a proof of the veracity of the model, in addition to giving restrictions on the Yukawas.

>>>>

% \subsection{Scalar mass spectrum}
%\label{sec:ScaMassSpect}
To establish the scalar spectrum we expand the scalar fields as
%
\begin{align*}
  H = \begin{pmatrix}G^+ \\ \frac{1}{\sqrt{2}} (\varphi+v_H+iG) \end{pmatrix} \,,&\hspace{1cm}
  \eta = \begin{pmatrix}\eta^{+} \\ \frac{1}{\sqrt{2}}(\eta_R+i\eta_I) \end{pmatrix} \,,\\
  S = \frac{1}{\sqrt{2}} (S_R+v_{S}+iS_I)\,,&\hspace{1cm}\sigma = \frac{1}{\sqrt{2}} (\sigma_R+i\sigma_I),
\end{align*}
%
with $v= 246.22$ GeV.  
By minimizing the scalar potential we find that the mass-squared parameters $\mu_H^2$ and $\mu_S^2$
%
\begin{align*}
    \mu^{2}_{H} =& -\frac{1}{2}\left( \lambda_{H}  v^2 + \lambda_{1}  v_S\right)\,,  &
    \mu^{2}_{S} =& -\frac{1}{2}\left( \lambda_{S}  v^{2}_{S} + \lambda_{1}  v^2\right)\,.
\end{align*}
%
Of the original twelve scalar degrees of freedom in the model, $W^{\pm}$, $Z^{0}$ and $Z^{\prime}$ absorve four of them (the Goldstone bosons $G^{\pm}$,$G$ and $S_{I}$). There is a physical neutral CP-even Higgs state ($\varphi$), one scalar state ($S_{R}$), two CP-even scalar states ($\sigma_{R}$ and $\eta_{R}$), two CP-odd scalar states ($\sigma_{I}$ and $\eta_{I}$) and one charged scalar ($\eta^{\pm}$). The mass spectrum for the scalar sector is given by
\begin{align*}
    m_{\eta^{\pm}}^{2} &= \mu_{\eta}^{2} + \frac{1}{2} (\lambda_{3} \upsilon^{2}_{H} + \lambda_{5} \upsilon^{2}_{s} )\,, \\
    m_{H}^{2} &= \lambda_{H} \upsilon_{H}^{2} + \mathcal{O}()\,, \\
    m_{S}^{2} &= \lambda_{S} \upsilon_{S}^{2} + \mathcal{O}()\,.
\end{align*}
%
The neutral mass eigenstate are defined as
%
\begin{align*}
    \begin{pmatrix}\chi_{(R,I)_{1}} \\ \chi_{(R,I)_{2}} \end{pmatrix} =& \begin{pmatrix} c_{\theta_{(R,I)}} & -s_{\theta_{(R,I)}} \\ s_{\theta_{(R,I)}} & c_{\theta_{(R,I)}} \end{pmatrix} \begin{pmatrix}\sigma_{(R,I)} \\ \eta_{(R,I)} \end{pmatrix} \,,
\end{align*}
%
where $\tan2\theta_{(R,I)} = \pm2c/(b-a)$ and we defined $a = m^{2}_{\eta^{\pm}} + \frac{1}{2}\lambda_{7} \upsilon^{2}_{H}$, $b = \mu^{2}_{\sigma}+\frac{1}{2}(\lambda_{2} \upsilon^{2}_{H} + \lambda_{4} \upsilon^{2}_{S})$ and $c = \frac{1}{2}\lambda_8 v_{H} v_{S}$\,. The corresponding mass eigenvalues are
\begin{align*}
    m_{\chi_{R(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} + 4c^{2}} \right)\,, \\
     m_{\chi_{I(1,2)}}^{2} &= \frac{1}{2} \left(a + b \mp \sqrt{(a-b)^{2} - 4c^{2}} \right)\,,
\end{align*}

On the other hand, we assume that the heavy Majorana fermions are already in the diagonal basis defined as $
\begin{pmatrix}
  \psi_1 &
  \psi_2 &
  \psi_3
\end{pmatrix}^{\operatorname{T}}
$, with masses $M_{\psi_a}$ ($a=1,2,3$).
\section{Neutrinos}
\label{sec:Neutrinos}
%
\begin{figure}
\centering
\includegraphics[scale=0.75]{Neutrino_Loop.pdf}
\caption{One-loop generation of Dirac neutrino mass}
\label{fig:zee}
\end{figure}
%
 The Dirac neutrino mass matrix in this model is generated at one loop-level according to the diagram in  fig.~\ref{fig:zee}. The expression for $\mathcal{M}_{\nu}$ is given by~\cite{Reig:2018mdk}
%
\begin{align}
(\mathcal{M}_{\nu})_{ij} = \frac{\lambda_8 v_S v}{32 \pi^{2}} \sum_{a=1}^{3} \frac{h_{i a} M_{\psi_{a}}y^{\dagger}_{a j}} {m_{\chi_{R_2}}^{2}-m_{\chi_{R_1}}^{2}} \left[ F\left( \frac{m_{\chi_{R_2}}^{2}}{M_{\psi_{a}}^{2}} \right) - F\left( \frac{m_{\chi_{R_1}}^{2}}{M_{\psi_{a}}^{2}} \right) \right] + (R \to I)\,,
\end{align}
%
%check the following formula:
where $F(x) =x \log x/(x-1)$ is the loop function, $M_{\psi_{a}}$ is
the fermion mass and $\chi_{R_{(1,2)}}$ ( $\chi_{I_{(1,2)}}$ ) are the
two CP-even (CP-odd) mass eigenstates.
To obtain neutrino masses it is necessary that $\mu \neq 0$. In this
way, in addition to the suppression from the seesaw mechanism, we can
have further suppression in the neutrino mass matrix for small values
of $\mu$.
The structure of the effective neutrino mass matrix is given by the
product $(M_{\nu})_{ij} \propto h_{i a} y_{a j}$, which is similar to
the structure of the neutrino mass matrix for the tree-level seesaw
mechanism for Dirac neutrinos~\cite{Chulia:2016ngi}.

If only one
fermion $\psi_R$ is added, we will have two massless neutrinos, which
does not agree with the current data obtained from the neutrino
experiments~\cite{deSalas:2017kay}. In our case, we will assume the
existence of three fermions, generating Dirac scotogenic masses for
the two neutrinos ($\nu_{k}$ is massless due to the assignment of
charges). If we consider the case in which $m_{\eta^{\pm}}^2 = \mu_{\sigma} \gg \frac{1}{2} \lambda_8 v v_S$ and $\lambda_7$, $\lambda_2$, $\lambda_4 \gg 1$. Taking $m_{\chi_{R_{2}}}-m_{\chi_{R_{1}}} = \lambda_8 v v_{S}$ and $m_{\chi_{R_{2}}}+m_{\chi_{R_{1}}} = 2 m_{\eta^{\pm}}^{2}$ one finds
%
\begin{align*}
(\mathcal{M}_{\nu})_{ij} = \frac{\lambda_8 v_S v}{16 \pi^{2}} \sum_{a=1}^{3} \frac{h_{i a} M_{\psi_{a}}y^{\dagger}_{a j}} {m_{\eta^{\pm}}^{2}-M_{\psi_{a}}^{2}} \left[1 - \frac{M_{\psi_{a}}^{2}}{m_{\eta^{\pm}}^{2}-M_{\psi_{a}}^{2}} \log \left( \frac{m_{\eta^{\pm}}^{2}}{M_{\psi_{a}}^{2}} \right)\right]\,,
\end{align*}
%
and if we take $m_{\eta^{\pm}}^{2} \gg M_{\psi_{a}}^{2}$ one has 
%
\begin{align}
(\mathcal{M}_{\nu})_{ij} & = \frac{\lambda_8 v_S v}{16 \pi^{2}} \sum_{a=1}^{3} \frac{h_{i a} M_{\psi_{a}}y^{\dagger}_{a j}} {m_{\eta^{\pm}}^{2}}\,, \\
& \sim 0.006~\text{eV} \left( \frac{\lambda_8}{1 \text{GeV}}\right) \left( \frac{v_{S}}{10 \text{TeV}}\right) \left( \frac{M_{\psi_{a}}}{1 \text{TeV}}\right) \left( \frac{50 \text{TeV}}{m_{\eta^{\pm}}}\right)^{2} \left( \frac{h_{i a} y_{a j}^{\dagger}}{10^{-6}}\right)\,. \nonumber
\end{align}
%
The mass matrix for Dirac neutrinos is diagonalized by
\begin{align*}
(U^{\nu}_R)^{\dagger} \mathcal{M}_{\nu} U^{\nu}_L = \mathcal{M}_{\nu}^{\text{diag}} = \text{diag}(m_1, m_2, m_3)\,,
\end{align*}
where $U^{(\nu)}_R$ and $U^{(\nu)}_L$ are unitary $3 \times 3$
matrices. The leptonic charged currents with non-massless neutrinos
are parametrized in terms of the unitary mass
\begin{align*}
U_{\text{PMNS}} = \left( U^{\ell}_L \right)^{\dagger} U^{\nu}_L\,,
\end{align*}
where is the left diagonalization matrix of the charged leptons.
As we already are in the diagonal basis after the calculation of the
one-loop neutrino masses, we can set
$U_{L}^{\ell} = \mathds{1}$. Moreover, we assume that $U_{R}^{\nu} = \mathds{1}$.
Therefore, we can express the Dirac neutrino mass matrix in the
following way
\begin{align}
(\mathcal{M}_{\nu})_{ij} = (U_{\text{PMNS}})_{ij} (\mathcal{M}_{\nu}^{\text{diag}})_{j}\,,
\end{align}
where the matrix $U_{\text{PMNS}}$ is expressed in the standard parameterization as
\begin{align*}
U_{\text{PMNS}} = \begin{pmatrix}
    1 & 0		& 0 \\
    0 & c_{23}  & s_{23} \\
    0 & -s_{23} & c_{23}
\end{pmatrix}
\begin{pmatrix}
    c_{13}  &  0 & s_{13}e^{-i\delta} \\
    0 		&  1 & 0 \\
    -s_{13}e^{i\delta} &  0 & c_{13}
\end{pmatrix}
\begin{pmatrix}
    c_{12}  & s_{12} & 0 \\
    -s_{12} & c_{12} & 0 \\
    0		& 0		 & 1
\end{pmatrix}\,,
\end{align*}
where $s_{ij} = \sin \theta_{ij}$ and $c_{ij} = \cos \theta_{ij}$ and $\theta_{ij}$ are neutrino mixing angles and $\delta$ is CP-violating phase.

@@

For $i \neq j$ we have CLFV processes, these processes are severely restricted by the current bounds , which set limits to the Yukawas. The most restrictive experimental bounds is decay $\mu \to e \gamma$, which is several orders of magnitude greater than radiative decays $\tau \to~l\gamma$, with $l = e, \mu$. These restrictions are given by Br$(\mu \to e\gamma)$ \textless $5.7 \times 10^{-13} $~\cite{Adam:2013mnn}, Br$(\tau \to e\gamma) $\textless$ 3.3 \times 10^{-8}$ and Br$(\tau \to~\mu\gamma) $\textless$ 4.4 \times 10^{-8}$~\cite{Aubert:2009ag, Bona:2007qt, Miyazaki:2012mx}.

In the limit $m_{\eta^{\pm}}^{2} \gg M_{\psi_{a}}^{2}$, $t_{\alpha} \to 0$ the decay width reads
%
\begin{align}
 \Gamma(l^{-}_{i} \to l^{-}_{j} \gamma) = \frac{m^{3}_{i}}{16 \pi} \left| \sum_{\alpha} h^{\alpha}_{i} m_{i} h^{\alpha}_{j} \frac{i e}{16 \pi M^{2}_{\eta^{-}}} \left[ \frac{1}{12} \right] \right|^{2}\,, 
\end{align}
%
using the current experimental constraint, can be fullfilled provided
%
\begin{align*}
    \left( \sum_{\alpha} \frac{h^{}_{i \alpha} h_{\alpha j}^{*}}{M^{2}_{\eta^{-}}} \right)^{2} \leq 5.7 \times 10^{-13} \frac{768 \pi G_{F}^{2}}{\alpha_{EM}}
\end{align*}
%
which leads to a very slight requirement,
\begin{align*}
    \left| \sum_{\alpha}h^{}_{i \alpha} h_{\alpha j}^{*}\left(\frac{50\text{TeV}}{M^{2}_{\eta^{-}}} \right)^{2} \right| \leq 12.65
\end{align*}
%
For $i = j$ we have processes that contribute to the leptonic magnetic dipole moment. In the case of the anomalous muon magnetic dipole moment, the experimental measurements does not agree with the values predicted by the SM~\cite{Lindner:2016bgg}. This can be a proof of the veracity of the model, in addition to giving restrictions on the Yukawas.
